\Large \bf 6.824 Final Project

### 2014-05-12 writeup/paper.tex
github link
Spell check

\section{Design} \label{sec:design} The design is based upon the code
developed throughout the semester for lab 4.  There are therefore
three basic components running on each server: Paxos processes to
reach consenus for instances, a shard master to determine which groups
are responsible for various shards of the database, and a shard
key/value storage process to execute client Put/Get requests.  The
Paxos instances are used to create a consistent log of operations for
both shard reconfiguration and client operations.

@@

\subsection{Paxos Optimizations}
In order to improve the performance of Paxos, we implemented the
Multi-Paxos leader selection optimization. This optimization solved
the dueling leaders problem, preventing us from hanging forever in a
loop between conflicting leaders, and also allowed us to reduce the
number of RPC calls that were necessary to come to agreement, by
allowing us to skip the Prepare phase of the Paxos protocol for a
leader that was constant between instances. The main issue with the
Multi-Paxos leader selection is ensuring that leadership information
is propogated through the system and ensuring that we never agree as
part of two separate partitions when the Prepare phase is
skipped. Unfortunately, we discovered a number of implementation bugs
with our default Paxos implementation, which we ended up fixing to end
up with a fairly reliable Multi-Paxos optimized system that had a
selected leader and skipped unnecessary prepare phases. The results
were that the number of required RPCs was reduced by about 33\%, and
that there were reduced bandwidth requirements due to the singular
leadership. There was an issue that a single machine became the
primary proposer, potentially limiting it based on disk and processor
speeds, however this doesn't appear to be an issue at the anticipated
use cases.

\subsection{Persistent Storage}
In order to tolerate servers restarting and to handle databases that
cannot fit in memory, all three components of the system store state
and data persistently on disk.  To interface with the disk, a Go
wrapper for the high-performance C++ key/value libray LevelDB was
utilized.  An outline of what each system component stores on disk is
presented below.
\begin{itemize}
\item The Paxos processes write the highest accepted value, the
  highest accepted proposal number, and the highest acknowledged
  prepare request to disk for each instance.  This ensures that the
  protocol will operate correctly even if a server restarts \cite{paxos} (for
  example, a decided value can never be changed).  In addition to this
  state, the Paxos processes persistently store the done responses
  received from peers and the number of the maximum known instance
  (see the recovery section below).

@@

\begin{table}
\begin{tabular}{|l|l|l|l|}
\hline
\multicolumn{4}{|c|}{\cellcolor[HTML]{C0C0C0}{\color[HTML]{000000} \textbf{Recovery Performance for 100MB Database}}}                           
\\ \hline \textbf{Value Size (KB)} &  \textbf{Number of Entries} & \textbf{Time Reading Database (s)} & \textbf{Recovery Time (s)} 
\\ \hline 5  & 20352 & 464.7 & ?
\\ \hline 50  & 2046 & 30.8 & 37.2
\\ \hline 100  & 1023 & 15.9 & 21.1
\\ \hline 150  & 682 & 10.2 & 15.8
\\ \hline 250  & 409 & 6.3 & 12.4
\\ \hline 500  & 204 & 3.5 & 10.6
\\ \hline 1024  & 99 & 1.88 & 8.3
\\ \hline 1500  & 68 & 1.4 & ?
\\ \hline
\end{tabular}
\end{table}

>>>>

Source code for this project is accessible at \url{https://github.com/pranjalv123/mexos}.

\section{Design} \label{sec:design} The design is based upon the code
developed throughout the semester for lab 4.  There are therefore
three basic components running on each server: Paxos processes to
reach consensus for instances, a shard master to determine which groups
are responsible for various shards of the database, and a shard
key/value storage process to execute client Put/Get requests.  The
Paxos instances are used to create a consistent log of operations for
both shard reconfiguration and client operations.

@@

\subsection{Paxos Optimizations}
In order to improve the performance of Paxos, we implemented the
Multi-Paxos leader selection optimization. This optimization solved
the dueling leaders problem, preventing us from hanging forever in a
loop between conflicting leaders, and also allowed us to reduce the
number of RPC calls that were necessary to come to agreement, by
allowing us to skip the Prepare phase of the Paxos protocol for a
leader that was constant between instances. The main issue with the
Multi-Paxos leader selection is ensuring that leadership information
is propagated through the system and ensuring that we never agree as
part of two separate partitions when the Prepare phase is
skipped. Unfortunately, we discovered a number of implementation bugs
with our default Paxos implementation, which we ended up fixing to end
up with a fairly reliable Multi-Paxos optimized system that had a
selected leader and skipped unnecessary prepare phases. The results
were that the number of required RPCs was reduced by about 33\%, and
that there were reduced bandwidth requirements due to the singular
leadership. There was an issue that a single machine became the
primary proposer, potentially limiting it based on disk and processor
speeds, however this doesn't appear to be an issue at the anticipated
use cases.

\subsection{Persistent Storage}
In order to tolerate servers restarting and to handle databases that
cannot fit in memory, all three components of the system store state
and data persistently on disk.  To interface with the disk, a Go
wrapper for the high-performance C++ key/value library LevelDB was
utilized.  An outline of what each system component stores on disk is
presented below.
\begin{itemize}
\item The Paxos processes write the highest accepted value, the
  highest accepted proposal number, and the highest acknowledged
  prepare request to disk for each instance.  This ensures that the
  protocol will operate correctly even if a server restarts \cite{paxos} (for
  example, a decided value can never be changed).  In addition to this
  state, the Paxos processes persistently store the done responses
  received from peers and the number of the maximum known instance
  (see the recovery section below).

@@

\begin{table}
\begin{tabular}{|l|l|l|l|}
\hline
\multicolumn{4}{|c|}{\cellcolor[HTML]{C0C0C0}{\color[HTML]{000000} \textbf{Recovery Performance for 100MB Database}}}                           
\\ \hline \textbf{Value Size (KB)} &  \textbf{Number of Entries} & \textbf{Time Reading Database (s)} & \textbf{Recovery Time (s)} 
\\ \hline 5  & 20352 & 464.7 & ~473
\\ \hline 50  & 2046 & 30.8 & 37.2
\\ \hline 100  & 1023 & 15.9 & 21.1
\\ \hline 150  & 682 & 10.2 & 15.8
\\ \hline 250  & 409 & 6.3 & 12.4
\\ \hline 500  & 204 & 3.5 & 10.6
\\ \hline 1024  & 99 & 1.88 & 8.3
\\ \hline 1500  & 68 & 1.4 & 
\\ \hline
\end{tabular}
\end{table}
